\section{Theory to Implementation}
\label{sec:implementation}

This section maps the formal model of \Cref{sec:model,sec:pricing,sec:offchain}
to the deployed smart contract architecture on Base Sepolia (L2).

\subsection{Contract architecture}

The core BPE protocol is realised by eight Solidity contracts,
with three additional thermodynamic contracts (\Cref{sec:thermodynamic}):

\begin{table}[h]
\centering
\small
\begin{tabular}{lll}
\toprule
\textbf{Contract} & \textbf{Formal object} & \textbf{Function} \\
\midrule
\texttt{CapacityRegistry} & $\Craw_s, \Csmooth_s$ & Stores raw + EWMA-smoothed capacity \\
\texttt{StakeManager} & $\sqrt{\cdot}$ cap & Concave staking, Sybil resistance \\
\texttt{BackpressurePool} & $\mu^*_{s,\tau}$ & GDA wrapper, Boltzmann rebalancing \\
\texttt{CompletionTracker} & Dual-signed receipts & Verifies task completions \\
\texttt{PricingCurve} & $p_d(t)$ & EIP-1559-style dynamic pricing \\
\texttt{EscrowBuffer} & Overflow buffer + $P$ & Holds payments; emits osmotic pressure \\
\texttt{OffchainAggregator} & BLS aggregation & Batch-verifies capacity attestations \\
\texttt{Pipeline} & Cross-pool routing & Composes flows; circuit breakers \\
\midrule
\texttt{TemperatureOracle} & $\tau(\sigma^2)$ & Variance-to-temperature mapping \\
\texttt{VirialMonitor} & $V = 2T/(S+E)$ & Capital equilibrium + adaptive $\delta$ \\
\texttt{SystemStateEmitter} & Phase classifier & Aggregates $\tau, V, P$ into snapshots \\
\bottomrule
\end{tabular}
\caption{Mapping from formal model to deployed contracts.}
\label{tab:contract-map}
\end{table}

\subsection{How theory maps to code}

\paragraph{Capacity smoothing (\Cref{sec:model}).}
The EWMA update $\Csmooth_s \leftarrow \alpha \cdot \Craw_s + (1 - \alpha) \cdot \Csmooth_s$
is computed in \texttt{CapacityRegistry.updateCapacity()} using fixed-point arithmetic
(18 decimals). The smoothing parameter $\alpha = 0.3$ is configurable per deployment.

\paragraph{Backpressure allocation (\Cref{sec:throughput}).}
The allocation policy $\mu^*_{s,\tau} \propto \Csmooth_s \cdot \sqrt{\text{stake}_s}$
is implemented in \texttt{BackpressurePool.rebalance()}, which updates Superfluid GDA
member units. Unit updates are atomic: a single transaction reads current capacities
from the registry, computes proportional shares, and sets GDA units.

\paragraph{Dynamic pricing (\Cref{sec:pricing}).}
The per-dimension base fee adjustment
$b_d(t+1) = b_d(t) \cdot (1 + \Delta \cdot (u_d - u^*_d))$
runs in \texttt{PricingCurve.updateBaseFee()} on each capacity update. The
target utilisation $u^* = 0.5$ and maximum adjustment $\Delta = 0.125$ are
the same values used in EIP-1559. The effective price a source pays is
$\sum_d C_{s,d} \cdot b_d(t)$.

\paragraph{Off-chain aggregation (\Cref{sec:offchain}).}
\texttt{OffchainAggregator} verifies a BLS aggregate signature over a batch of
capacity attestations in a single \texttt{ecrecover}-equivalent operation. In
benchmarks (Base Sepolia, 50 attestations per batch), this reduces gas cost by
83.5\% compared to verifying each attestation individually.

\paragraph{Completion verification (\Cref{sec:security}).}
\texttt{CompletionTracker.submitCompletion()} requires two EIP-712 signatures:
one from the agent (sink) and one from the requester (source). A completion
missing either signature is rejected. Completed tasks update the agent's
effective capacity and, when a QualityOracle is attached, feed into quality
scoring.

\subsection{Deployment}

All 32 contracts (8 core + 5 v2 extensions + 3 thermodynamic + 9 research + 7 platform/routing modules)
are deployed and verified on Base Sepolia. The TypeScript SDK (\texttt{@puraxyz/sdk})
provides 23 action modules covering registration, staking, capacity updates, stream
management, pipeline composition, and thermodynamic monitoring. The test suite
contains 319 passing tests spanning unit tests, integration tests, and gas benchmarks.

Source code, deployment addresses, and ABI exports are available at
\url{https://github.com/puraxyz/puraxyz}.
